\bf On Devon Allen's Disqualification at the 2022 World Track and Field Championships

### 2023-11-20 Code/RankComparison.Rmd
making sure everything is consistent

The code below performs two clustered Wilcox tests, the first one without using
exact = True, and the second one with exact = True and 1,000,000 random
permutations. The tests compare the times and group by the Stage of the
competition which in the data is coded as 1 if it was a national championship
and 2 if it was the world championships. The tests cluster by athlete, of which
for each dataset an integer from 1-11 is used (the values are not significant).
For more information about the "ds" method, please look at the Rank-based
Comparison subsection of the Methods section of the paper or at this link here:
https://www.jstatsoft.org/article/view/v096i06.  The second test also includes
exact = TRUE which is used when sample sizes are small (as they are in this
case), additionally the number of random permutations (B) is set to be 1,000,000.

@@

The code below reads in the data comparing the 2019 and 2022 World Championships.
```{r}
old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(old_clusrank) <- c("Time", "Athlete", "Year")
#View(comp)
```

@@

The code below reads in data comparing the 2022 and 2023 World Championships
```{r}
new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
#"C:\Users\ofior\Documents\Thesis\Data\2019vs2022.csv"
colnames(new_clusrank) <- c("Time", "Athlete", "Year")
#View(comp)
```

@@

This is seeing what would happen if we remove prelims, the sample sizes in the
test become very small: 26 observations and 9 clusters.
```{r}
#nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
#colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
```

>>>>

The code below performs two clustered Wilcox tests, the first one without using
exact = TRUE, and the second one with exact = TRUE and 1,000,000 random
permutations. The tests compare the times and group by the Stage of the
competition, which in the data is coded as 1 if it was a national championship
and 2 if it was the world championships. The tests cluster by athlete, of which
for each dataset an integer from 1-11 is used (the values are arbitrary).
For more information about the "ds" method, please look at the Rank-based
Comparison subsection of the Methods section of the paper or at this link here:
https://www.jstatsoft.org/article/view/v096i06.

@@

The code below reads in the data comparing the 2019 and 2022 World Championships.
```{r}
old_clusrank <- read.csv("../Data/Clusrank2019vs2022.csv")
colnames(old_clusrank) <- c("Time", "Athlete", "Year")
```

@@

The code below reads in data comparing the 2022 and 2023 World Championships
```{r}
new_clusrank <- read.csv("../Data/Clusrank2022vs2023.csv")
colnames(new_clusrank) <- c("Time", "Athlete", "Year")
```

@@

The following code could be used to see what happens if we remove prelims. 
The sample sizes in the test become very small: 26 observations and 9 clusters.
```{r}
#nations_noprelim <- read.csv("../Data/Clusrank2019vs2022NoPrelims.csv")
#colnames(nations_noprelim) <- c("Time", "Athlete", "Year")
```

### 2023-11-20 Manuscript/manuscript.tex
making sure everything is consistent

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $0.028$ & $0.003$ \\
   2022 vs 2023 & $0.011$ & $0.003$ \\
   2022 National vs International & $2\cdot10^{-6}$ & $0.001$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation and asymptotic rank-based tests for the three comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparison produced lower p-values than the 2019 versus 2022 comparison, with the
national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
This is a highly significant test result and shows that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior. For the 2019 versus 2022
same athlete comparison with permutation test, the p-value was 
$0.028$, which is significant at level $0.05$.  The result of
comparing 2022 and 2023 was very similar (but not exactly the same) to the above
result, as the p-value from the permutation test was $0.011$.
Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed at other
competitions.

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three models fitted to the
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
    \midrule
    Excluding 2022 &
      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.880$ & 0.136\\[1ex]
    Including 2022 &
    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.150\\
    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.880$ & 0.134\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than threshold 0.08,
  0.09, and 0.10 seconds based on the two effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\ 
   \midrule
   Excluding 2022 & $3.35\cdot10^{-5}$ & $4.45\cdot10^{-4}$ &  $3.54\cdot10^{-3}$  \\ 
   Including 2022 & $4.20\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.34\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}

>>>>

\begin{table}
  \centering
  \caption{2019 vs 2022 compares data from the same athletes who competed at the
  2019 and 2022 World Track and Field Championships. 2022 vs 2023 compares data
  from the same athletes who competed at the 2022 and 2023 World Track and Field
  Championships. National vs International compares data from the same athletes
  who competed at 2022 national-level championships and the 2022 World Track and
  Field Championships.}
  \begin{tabular}{c c c c c} 
   \toprule
   Comparison & Permutation & Asymptotic \\ 
   \midrule
   2019 vs 2022 & $0.028$ & $0.003$ \\
   2022 vs 2023 & $0.011$ & $0.003$ \\
   2022 National vs International & $<0.001$ & $0.002$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Clusrankresults}
\end{table}

Table~\ref{tab:Clusrankresults} shows the results generated from both the 
permutation and asymptotic rank-based tests for the three comparisons, 
as discussed previously in Section~\ref{sec:rank}. The national versus international
comparisons produced lower p-values than the 2019 versus 2022 and 2022 versus 2023comparisons.%, with the
%national vs international permutation test producing a p-value of $2\cdot10^{-6}$.
The highly significant permutation test result provides substantial evidence 
that athletes who competed at the
2022 World Track and Field Championships were on average, significantly faster
at reacting than they were just a few months prior. For the 2019 versus 2022
same athlete comparison with permutation test, the p-value was 
$0.028$, which is significant at level $0.05$.  The result of
comparing 2022 and 2023 was very similar (but not exactly the same) to the above
result, as the p-value from the permutation test was $0.011$.
Taken together, these three
tests help show that athletes competing at the 2022 World Track and Field
Championships had faster reaction times than when they had competed at other
competitions.

@@

\begin{table}
  \centering
  \caption{AIC and estimated parameters for the three models fitted to the
    data.}
  \label{tab:Gamma_parameters}
  \begin{tabular}{c c c c c c c}
    \toprule
    Data set & Model & AIC & $\sigma_h$ & $\sigma_v$ & $\alpha$ & $\sqrt{\phi}$ \\
    \midrule
    Excluding 2022 &
      Venue Effect Only & $-1862.2$ &   & 0.048 & $-1.857$ & 0.152\\
     & Heat Effect Only & $-1910.9$ & 0.069 &  & $-1.877$ & 0.136\\
     & Venue and Heat Effect & $-1912.9$ & 0.058 & 0.033 & $-1.868$ & 0.136\\[1ex]
    Including 2022 &
    Venue Effect Only & $-2023.5$ &   & 0.054 & $-1.871$ & 0.149\\
    & Heat Effect Only & $-2070.7$ & 0.073 &  & $-1.890$ & 0.135\\
    & Venue and Heat Effect & $-2075.7$ & 0.057 & 0.040 & $-1.881$ & 0.134\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}
  \centering
  \caption{Probabilities of observing reaction times less than threshold 0.08,
  0.09, and 0.10 seconds based on the two effect model.}
  \begin{tabular}{c c c c} 
   \toprule
   Data Set & Threshold 0.08 & Threshold 0.09 & Threshold 0.10  \\ 
   \midrule
   Excluding 2022 & $3.31\cdot10^{-5}$ & $4.46\cdot10^{-4}$ &  $3.52\cdot10^{-3}$  \\ 
   Including 2022 & $4.16\cdot10^{-5}$ & $5.70\cdot10^{-4}$ & $4.38\cdot10^{-3}$ \\
   \bottomrule
  \end{tabular}
  \label{tab:Sim_probability}
\end{table}
